---
title: "Business Intelligence Project"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |     |
|----------------------------------------------|-----|
| **Student ID Number**                        | 104135 |
| **Student Name**                             | Beasts |
| **BBIT 4.2 Group**                           | Beasts |
| **BI Project Group Name/ID (if applicable)** | group B (beats) |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

# Dataset loader

```{r}
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

data(PimaIndiansDiabetes)
```

# Split the dataset

```{r}
str(PimaIndiansDiabetes)
```

**Description:** 40% of the original data will be used to train the model and 60% of the original data will be used to test the model.

```{r}
library(caret)

train_index <- createDataPartition(PimaIndiansDiabetes$diabetes,
                                   p = 0.40,
                                   list = FALSE)
PimaIndiansDiabetes_train <- PimaIndiansDiabetes[train_index, ]
PimaIndiansDiabetes_test <- PimaIndiansDiabetes[-train_index, ]
```

## Train a Naive Bayes classifier using the training dataset

### naiveBayes() function in the e1071 package

```{r}
PimaIndiansDiabetes_model_nb_e1071 <-
  e1071::naiveBayes(diabetes ~ .,
                    data = PimaIndiansDiabetes_train)
```

### naiveBayes() function in the caret package

```{r}
PimaIndiansDiabetes_model_nb_caret <- # nolint
  caret::train(diabetes ~ ., data =
               PimaIndiansDiabetes_train[, c("pregnant", "glucose", "pressure",
                    "triceps", "insulin","mass", "pedigree","age","diabetes")],
               method = "naive_bayes")
```

## Test the trained model using the testing dataset

### Test the trained e1071 Naive Bayes model using the testing dataset

```{r}
predictions_nb_e1071 <-
  predict(PimaIndiansDiabetes_model_nb_e1071,
          PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
                        "triceps", "insulin","mass","pedigree","age","diabetes")])
```

### Test the trained caret Naive Bayes model using the testing dataset

```{r}
predictions_nb_caret <-
  predict(PimaIndiansDiabetes_model_nb_caret,
          PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
                        "triceps", "insulin","mass","pedigree","age","diabetes")])
```

## View the Results

### e1071 Naive Bayes model and test results using a confusion matrix

```{r}
print(predictions_nb_e1071)
caret::confusionMatrix(predictions_nb_e1071,
                       PimaIndiansDiabetes_test[, c("pregnant", "glucose",                               "pressure","triceps", "insulin","mass","pedigree","age",
                                                    "diabetes")]$diabetes)
plot(table(predictions_nb_e1071,
           PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
           "triceps", "insulin","mass","pedigree","age","diabetes")]$diabetes))
```
